---
title: "PSTAT 131 Final Project: Model comparison for predicting the salary of a data science job"
author: "Nicholas Axl Andrian"
date: "2023-11-23"
output: html_document
---

Dataset used: https://www.kaggle.com/datasets/arnabchaki/data-science-salaries-2023

In this project, we will be fitting several different machine learning algorithms to find out which method of prediction is the most accurate in getting the predicted salary.

About the dataset's variables (excerpt from the kaggle site)

+ work_year: The year the salary was paid.
+ experience_level: The experience level in the job during the year
+ employment_type: The type of employment for the role
+ job_title: The role worked in during the year.
+ salary: The total gross salary amount paid.
+ salary_currency: The currency of the salary paid as an ISO 4217 currency code.
+ salaryinusd: The salary in USD
+ employee_residence: Employee's primary country of residence in during the work + year as an ISO 3166 country code.
+ remote_ratio: The overall amount of work done remotely
+ company_location: The country of the employer's main office or contracting branch
+ company_size: The median number of people that worked for the company during the year


# Part 1: Exploratory Data Analysis

```{r setup, include=FALSE}
library(dplyr)
library(randomForest)
library(gbm)
library(ISLR)
library(tree)
library(tidyverse)
```

#### Loading the dataset
```{r}
salaries <- read.csv("ds_salaries.csv")
```
#### Checking the structure of the dataset
```{r}
head(salaries)
str(salaries)
```
#### Checking the summary of the dataset
```{r}
summary(salaries)
```
#### Checking for null values
```{r}
colSums(is.na(salaries))
```
Fortunately, we have no null values so imputing is not required




